---
title: "Reproducing Open Science Research 3"
df-print: kable
execute: 
  echo: true
  warning: false
---

**October 06, 10am to 11am. [Register](https://arizona.zoom.us/meeting/register/tZIod-6urjksHdyNtrnlOfB49lh0hKgOLFQ5)**

For this tutorial, we will be replicating some of the analysis in [*Chappell, W. & Kanwit, M. (2021). Do Learners Connect Sociophonetic Variation with Regional and Social Characteristics? The Case of L2 Perception of Spanish Aspiration. Studies in Second Language Acquisition. 44(1). 1–25.*](https://scholar.google.com/scholar?hl=en&as_sdt=0%2C3&q=Chappell%2C+W.+%26+Kanwit%2C+M.+%282021%29.+Do+Learners+Connect+Sociophonetic+Variation+with+Regional+and+Social+Characteristics%3F+The+Case+of+L2+Perception+of+Spanish+Aspiration.+Studies+in+Second+Language+Acquisition.+44%281%29.+1%E2%80%9325.&btnG=#:~:text=include%20citations-,%5BPDF%5D%20cambridge.org,-Do%20learners%20connect)

Data can be found at <https://www.iris-database.org/details/MQbI5-rz7z3>

# Overview of study

* Matched-guise test targeting coda /s/ in Spanish
* Question: Do L2 Spanish learners identify native speakers' social characteristics based on phonetic variants

# Data

Our first step it to load the `tidyverse` library (RStudio will prompt you to install it).

```{r}
#| echo: true
#| eval: true
#| warning: false
#| message: false
# load library for data analysis
library(tidyverse)
```

Now we can read the data in. Remember to download the file and place it in a directory called `data` in your project. 

```{r}
#| echo: true
#| eval: true
#| warning: false
#| message: false
l2_span_perception <- read_csv("data/L2PerceptionsofAspirationAnonymized.csv")
```

## Participants

> Seventy-six language learners

We have multiple observations per participants, so to get to number of participants we need to use `distinct()`

```{r}
#| echo: true
#| eval: true
l2_span_perception %>% 
  distinct(Participant) %>% 
  nrow()
```

Table 1 recreation (page 193):

```{r}
#| echo: true
#| eval: true
l2_span_perception %>% 
  distinct(Participant, Gender) %>% 
  count(Gender)
```

```{r}
#| echo: true
#| eval: true
l2_span_perception %>% 
  distinct(Participant, Age) %>% 
  summarize(min_age = min(Age),
            max_age = max(Age),
            mean_age = mean(Age),
            median_age = median(Age))
```


```{r}
#| echo: true
#| eval: true
l2_span_perception %>% 
  distinct(Participant, Education) %>% 
  count(Education, sort = TRUE)
```


# Research questions

* Are L2-Spanish language learners bale to perceive the social meanings of coda /s/ in their L2?

> If so, what social properties will L2 Learners associate with coda [h]?

```{r}
#| echo: true
#| eval: true
l2_span_perception %>% 
  ggplot(aes(x = VariantHeard,
             y = PerceivedAge)) +
  geom_boxplot()
```

```{r}
#| echo: true
#| eval: true
l2_span_perception %>% 
  group_by(VariantHeard) %>% 
  summarize(mean_age = mean(PerceivedAge)) %>% 
  ggplot(aes(x = VariantHeard,
             y = mean_age)) +
  geom_col()
```


```{r}
#| echo: true
#| eval: true
l2_span_perception %>% 
  group_by(VariantHeard, SpeakerType, PerceivedSpeakerOrigin) %>% 
  summarize(count = n()) %>% 
  mutate(total = sum(count),
         percent = round((count/total)*100, 2)) %>% 
  ggplot(aes(x = PerceivedSpeakerOrigin, y = SpeakerType,
             fill = count)) +
  geom_tile() +
  geom_label(aes(label = paste0(percent, "%")),
             fill = "white") +
  facet_wrap(~VariantHeard) +
  scale_fill_gradient(low = "lightgray" , high = "black")
```

# Factor Analysis

> "[...] we conducted an FA and used the Kaiser Rules to establish which properties should be combined and analyzed as joint factors in the model-construction procedure"

For the Factor Analysis we need a data frame with only the indices we will be running on the analysis.

```{r}
#| echo: true
#| eval: true
indices <- l2_span_perception %>% 
  select(Intelligent:Feminine)
```

Now we can run the FA (with 3 factors, since that's what the authors did).

```{r}
#| echo: true
#| eval: true
fa <- factanal(x = indices, factors = 3) 
fa
```

Then we inspect the loadings.

```{r}
#| echo: true
#| eval: true
fa$loadings
```


```{r}
#| echo: false
#| eval: false
loadings <- fa$loadings
dimnames(loadings)[[1]]
loadings[1:8]
```

> The FA motivated the creation of three combined factors: (a) a status factor (loading for intelligence and work ethic), (b) a confident Spanish-speaker factor (loading for Hispanicity, confidence, and good Spanish), and (c) a solidarity factor (loading for niceness and humility). As no other factors appeared to be correlated, they were explored independently.

# Regression models

> Mixed-effects regression models were then created using the lme4 (Bates et al., Reference Bates, Maechler, Bolker and Walker2017) and lmerTest (Kuznetsova et al., Reference Kuznetsova, Brockhoff and Christensen2016) packages in R (R Core Team, 2018), and individual models were fitted to the following dependent variables: (a) status (intelligence/work ethic), (b) confident Spanish speaker (Hispanicity/confidence/good Spanish), (c) solidarity (niceness/humility), (d) age, (e) femininity, and (f) perceived speaker origin. Treatment contrasts were used, and the random effects in each model included the listener and the presentation order of the stimuli.

> The independent variables tested in each model include variant ([s] or [h]), speaker type (Mexican or Puerto Rican), having taken a phonetics class (yes or no), most advanced Spanish class taken divided into four collapsed categories (elementary, intermediate low, intermediate high, and advanced), number of weeks spent studying abroad (continuous), experience abroad with an aspirating variety (yes or no), whether participants use Spanish regularly with NSs (yes or no), whether participants use Spanish regularly at work (yes or no), whether participants listen regularly to Spanish media (e.g., shows, podcasts, movies, music [yes or no]), listener age (continuous), and listener gender (man, woman, or other). 

## dependent variable: status (intelligence/work ethic)

```{r}
#| echo: true
#| eval: true
l2_span_perception <- l2_span_perception %>% 
  mutate(dep_status = Intelligent + Hardworking)
```

```{r}
#| echo: true
#| eval: true
library(lme4)
library(lmerTest)
model_status <- lmer(dep_status ~ VariantHeard + SpeakerType + PhoneticsClass +
                     MaxClassBinned + WeeksAbroad + AspirationContact +
                     UseWithNSs + UseAtWork + UseListeningToMedia + Age +
                     Gender + (1|Participant) + (1|AudioNumber),
                   data = l2_span_perception)

step_model_status <- step(model_status)
step_model_status
```

```{r}
model_status_2 <- lmer(dep_status ~ VariantHeard + PhoneticsClass +
                         VariantHeard:PhoneticsClass +
                     MaxClassBinned + (1|Participant) + (1|AudioNumber),
                   data = l2_span_perception)

summary(model_status_2)
anova(model_status_2)
```


```{r}
#| eval: false
#| echo: false
model_2_results <- summary(model_status_2)
# calculate McFadden's R-squared for model
with(summary(model_status_2), 1 - deviance/null.deviance)
```

```{r}
library(MuMIn)
r.squaredGLMM(model_status_2)
```


# speakers' perceived place of origin (Caribbean vs. other)

Table 2 replication

```{r}
#| echo: true
#| eval: true
l2_span_perception <- l2_span_perception %>% 
  mutate(origin_dep = if_else(PerceivedSpeakerOrigin == "Caribbean", 1, 0),
         VariantHeard = factor(VariantHeard, levels = c("s", "h")))

model_origin <- lmer(origin_dep ~ VariantHeard + PhoneticsClass +
                         VariantHeard:PhoneticsClass +
                     MaxClassBinned + (1|Participant) + (1|AudioNumber),
                   data = l2_span_perception)

summary(model_origin)
anova(model_origin)
r.squaredGLMM(model_origin)
```

> To clarify this rather complex relationship, a conditional inference tree is provided in Figure 2.

Oh no

```{r}
library(effects)
library(ggthemes)
effect("VariantHeard:PhoneticsClass", model_origin) %>% 
  data.frame() %>% 
  ggplot(aes(x = PhoneticsClass,
             y = fit,
             ymin = lower,
             ymax = upper,
             color = VariantHeard)) +
  geom_errorbar() +
  geom_label(aes(label = round(fit, 2))) +
  theme_linedraw() +
  scale_color_colorblind() +
  labs(title = "Probability (fit) of participant hearing /s/")
```

```{r}
library(effects)
library(ggthemes)
effect("MaxClassBinned", model_origin) %>% 
  data.frame() %>% 
  mutate(MaxClassBinned = factor(MaxClassBinned,
                                 levels = c("Elementary",
                                            "Intermediate Low",
                                            "Intermediate High",
                                            "Advanced"))) %>% 
  ggplot(aes(x = MaxClassBinned,
             y = fit,
             ymin = lower,
             ymax = upper)) +
  geom_errorbar() +
  geom_label(aes(label = round(fit, 2))) +
  theme_linedraw() +
  labs(title = "Probability (fit) of participant hearing /s/")
```

# Conditional Inference Trees

```{r}
library(partykit)
l2_span_perception <- l2_span_perception %>% 
  mutate(VariantHeard = factor(VariantHeard),
         PhoneticsClass = factor(PhoneticsClass),
         MaxClassBinned = factor(MaxClassBinned),
         PerceivedSpeakerOrigin = factor(PerceivedSpeakerOrigin))

ctree_model =  ctree(PerceivedSpeakerOrigin ~ VariantHeard + PhoneticsClass + MaxClassBinned, 
                     data = l2_span_perception)

plot(ctree_model)
```
