---
title: "Reproducing Open Science Research 3"
df-print: kable
---

**September 22, 10am to 11am. [Register](https://arizona.zoom.us/meeting/register/tZIod-6urjksHdyNtrnlOfB49lh0hKgOLFQ5)**

For this tutorial, we will be replicating some of the analysis in [*Chappell, W. & Kanwit, M. (2021). Do Learners Connect Sociophonetic Variation with Regional and Social Characteristics? The Case of L2 Perception of Spanish Aspiration. Studies in Second Language Acquisition. 44(1). 1–25.*](https://scholar.google.com/scholar?hl=en&as_sdt=0%2C3&q=Chappell%2C+W.+%26+Kanwit%2C+M.+%282021%29.+Do+Learners+Connect+Sociophonetic+Variation+with+Regional+and+Social+Characteristics%3F+The+Case+of+L2+Perception+of+Spanish+Aspiration.+Studies+in+Second+Language+Acquisition.+44%281%29.+1%E2%80%9325.&btnG=#:~:text=include%20citations-,%5BPDF%5D%20cambridge.org,-Do%20learners%20connect)

Data can be found at <https://www.iris-database.org/details/MQbI5-rz7z3>

# Overview of study

* Matched-guise test targeting coda /s/ in Spanish
* Question: Do L2 Spanish learners identify native speakers' social characteristics based on phonetic variants

# Data

Our first step it to load the `tidyverse` library (RStudio will prompt you to install it).

```{r}
#| echo: true
#| eval: true
#| warning: false
#| message: false
# load library for data analysis
library(tidyverse)
```

Now we can read the data in. Remember to download the file and place it in a directory called `data` in your project. 

```{r}
#| echo: true
#| eval: true
#| warning: false
#| message: false
l2_span_perception <- read_csv("data/L2PerceptionsofAspirationAnonymized.csv")
```

## Participants

> Seventy-six language learners

We have multiple observations per participants, so to get to number of participants we need to use `distinct()`

```{r}
#| echo: true
#| eval: true
l2_span_perception %>% 
  distinct(Participant) %>% 
  nrow()
```

Table 1 recreation (page 193):

```{r}
#| echo: true
#| eval: true
l2_span_perception %>% 
  distinct(Participant, Gender) %>% 
  count(Gender)
```

```{r}
#| echo: true
#| eval: true
l2_span_perception %>% 
  distinct(Participant, Age) %>% 
  summarize(min_age = min(Age),
            max_age = max(Age),
            mean_age = mean(Age),
            median_age = median(Age))
```


```{r}
#| echo: true
#| eval: true
l2_span_perception %>% 
  distinct(Participant, Education) %>% 
  count(Education, sort = TRUE)
```


# Research questions

* Are L2-Spanish language learners bale to perceive the social meanings of coda /s/ in their L2?

> If so, what social properties will L2 Learners associate with coda [h]?

```{r}
#| echo: true
#| eval: true
l2_span_perception %>% 
  ggplot(aes(x = VariantHeard,
             y = PerceivedAge)) +
  geom_boxplot()
```

```{r}
#| echo: true
#| eval: true
l2_span_perception %>% 
  group_by(VariantHeard) %>% 
  summarize(mean_age = mean(PerceivedAge)) %>% 
  ggplot(aes(x = VariantHeard,
             y = mean_age)) +
  geom_col()
```


```{r}
#| echo: true
#| eval: true
l2_span_perception %>% 
  group_by(VariantHeard, SpeakerType, PerceivedSpeakerOrigin) %>% 
  summarize(count = n()) %>% 
  mutate(total = sum(count),
         percent = round((count/total)*100, 2)) %>% 
  ggplot(aes(x = PerceivedSpeakerOrigin, y = SpeakerType,
             fill = count)) +
  geom_tile() +
  geom_label(aes(label = paste0(percent, "%")),
             fill = "white") +
  facet_wrap(~VariantHeard) +
  scale_fill_gradient(low = "lightgray" , high = "black")
```

# Factor Analysis

> "[...] we conducted an FA and used the Kaiser Rules to establish which properties should be combined and analyzed as joint factors in the model-construction procedure"

```{r}
#| echo: true
#| eval: true
indices <- l2_span_perception %>% 
  select(Intelligent:Feminine)

fa <- factanal(x = indices, factors = 3) 
fa
```

```{r}
#| echo: true
#| eval: true
loadings <- fa$loadings
dimnames(loadings)[[1]]
```


```{r}
loadings[1:8]
```

